---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
    echo=TRUE
---


## Loading and preprocessing the data
```{r}
library(ggplot2)
library(dplyr)
dat = read.csv("activity.csv")
dat$date=as.Date(x=dat$date,format="%Y-%m-%d")
dat = as.tbl(dat)
```


## What is mean total number of steps taken per day?
1) A histogram of the total number of steps taken each day
```{r Histogram_StepPerDay_original}
tmp=group_by(dat,date)%>%summarise(total=sum(steps,na.rm=T))
ggplot(tmp,aes(x=date,y=total))+geom_histogram(stat="identity")+ylab("Number of Steps")

StepMean = mean(tmp$total,na.rm=T)
StepMedian = median(tmp$total)
```

2) The mean of the total number of steps per day is `r StepMean` and the median is `r StepMedian`

## What is the average daily activity pattern?
1) Time series plot of the average number of steps per interval
```{r TimeSeries_StepPerInterval}
AvgStepPerInterval=group_by(dat,interval)%>%summarise(total=mean(steps,na.rm=T))
ggplot(AvgStepPerInterval,aes(x=interval,y=total))+geom_line()+ylab("Average Number of Steps")
MaxInterval = AvgStepPerInterval[which.max(AvgStepPerInterval$total),1]
```
2) The interval with the maximum average number of steps is `r MaxInterval`

## Imputing missing values

1) Missing values
```{r}
NumMissingValues = sum(is.na(dat$steps))
```
Number of missing values is `r NumMissingValues`

2,3) Used the mean of the 5-minute interval (rounded) to fill missing values, and create a new dataset
```{r}
ImputedDat=inner_join(dat,AvgStepPerInterval,"interval")
ImputedDat[is.na(ImputedDat$steps),1]=round(ImputedDat[is.na(ImputedDat$steps),4])
ImputedDat = select(ImputedDat,-total)
```

4) A histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r Histogram_StepPerDay_imputed}
tmp=group_by(ImputedDat,date)%>%summarise(total=sum(steps,na.rm=T))
ggplot(tmp,aes(x=date,y=total))+geom_histogram(stat="identity")+ylab("Number of Steps")

StepMeanImputed = mean(tmp$total)
StepMedianImputed = median(tmp$total)
```

The mean of the total number of steps per day using the imputed data is `r StepMeanImputed` and the median is `r StepMedianImputed`

The mean using the imputed data is much bigger than using the original data (a difference of `r abs(StepMean-StepMeanImputed)`)

The median is also different by `r abs(StepMedian-StepMedianImputed)`. Note that the median increased because most of the missing values are at earlier dates. 

## Are there differences in activity patterns between weekdays and weekends?
1- Adding a new factor indicating whether a given date is a weekday or weekend
```{r}
weekends = weekdays(ImputedDat$date,abbreviate=TRUE) %in% c("Sun","Sat")
ImputedDat = mutate(ImputedDat,DateType="weekday")
ImputedDat$DateType[weekends] = "weekend"
```

2-Compare the average number of steps during weekends and weekdays
```{r TimeSeries_Weekend_WeekDays}
tmp=group_by(ImputedDat,interval,DateType)%>%summarise(total=mean(steps,na.rm=T))
ggplot(tmp,aes(x=interval,y=total))+geom_line()+ylab("Average Number of Steps") + facet_wrap(~DateType,ncol=1)
```
